---
title: "Subpopulations Manuscript Statistical Analysis"
output:
  html_document:
    df_print: paged
---

This notebook contains all of the statistical analysis and outputs for the 
manuscript, "Differentially tolerant and metabolically distinct subpopulations 
in shaken batch cultures of Pseudomonas aeruginosa", submitted to NBJ Biofilms. 

There are three main experiments in this analysis:

[Growth Phase and Antibiotic Tolerance] 

[Antibiotic Tolerance in Fresh vs. Used Media] 

[Fluorescence of Planktonic and Aggregate Subpopulations] (#anchor)

# Growth Phase and Antibiotic Tolerance

Note: Experiments 1 and 2 were discarded due to experimental errors and inconsistencies with the method. 
Thus, the analysis in the paper only contains the data from experiments 3-6.   


##Statistical Analaysis

```{r, include=FALSE}
#load libraries 
library(ggplot2)
library(plyr)
library(emmeans)
library(multcomp)
library(multcompView)
library(psych)
library(knitr)
library(cowplot)
library(tidyr)
library(magick)


```


#Log Density 

```{r, echo=TRUE}
#########  Analysis - Growth vs Time ##############

########Log Density and descriptive statistics ############################################################

#Import data

antibiotics.ld <- read.delim("./Data/GrowthPhaseAntibiotics/antibiotics.ld.txt")

#manually remove NAs and extra columns 

antibiotics.ld<-antibiotics.ld[-c(53,239,240,241:249),c(1,2,3,5)]

#Set variables

antibiotics.ld$Time<-as.factor(antibiotics.ld$Time)
antibiotics.ld$Log.LD<-as.numeric(as.character(antibiotics.ld$Log.LD))

#only experiments 3-6 

antibiotics.ld.sub<-subset(antibiotics.ld,Experiment==3|Experiment==4|Experiment==5|Experiment==6)

##Import summary stats from excel file (im too lazy rn to figure out how to automatically generate) 
sum.stats<- read.delim("./Data/GrowthPhaseAntibiotics/MLD_and_SE_Exp3to6.txt")

sum.stats$Treatment <- as.factor(sum.stats$Treatment)
sum.stats$Time <- as.character(sum.stats$Time)

#Change factor names and reorder
sum.stats$Treatment<-revalue(sum.stats$Treatment,c("0"="Control","1"="Tobramaycin (100 μg/mL) ","2"="Colistin (25 μg/mL)","3"="Tobramycin (100 μg/mL)\nColistin (25 μg/mL)"))

#Colors: 
colors <- c("#d55e00", "#56b4e9", "#f0e442", "#cc79a7")

#LD plot 
library(ggplot2)

plt_LD<-
ggplot(data=sum.stats, aes(x=as.character(Time), y=MLD, fill=Treatment)) +
  geom_bar(stat="identity", position='dodge')+
  scale_x_discrete(limits=c("2","4","8","24","48"))+
  scale_y_continuous(limits=c(0,10),breaks=c(0,2,4,6,8,10))+
  scale_fill_manual(values = colors)+
  ggtitle("Mean Log Density vs. Time")+
  ylab("Mean Log Density (log[CFU/mL])")+
  xlab("Time(hr)")+
  theme(axis.title.x = element_text(size=14,face="bold"),
        axis.title.y = element_text(size=14,face="bold"),
        title=element_text(size=16,face="bold"),
        axis.text=element_text(size=14,face="bold"))+
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=MLD-SE, ymax=MLD+SE),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9))+
  facet_grid(~Treatment)+
  theme_classic()
        
plt_LD

ggsave("./figures/plt_LDvsTime.png", plt_LD, height = 5, width = 9, dpi = 300)



```

##Statistical analysis - Log Reduction

```{r, echo=TRUE}

# Log Reduction #

#Reading in the files

antibiotics.lr<- read.delim("./Data/GrowthPhaseAntibiotics/antibiotics.lr.txt")

#Remove NA's and extra columns

antibiotics.lr<-antibiotics.lr[1:60,1:4]

#Setting variables to factors

antibiotics.lr$Experiment<-factor(antibiotics.lr$Experiment)
antibiotics.lr$Treatment<-factor(antibiotics.lr$Treatment)
antibiotics.lr$Time<-factor(antibiotics.lr$Time)

#make the subset - Experiments 3-6 because the first three had strange results, due to technician error.#
antibiotics.lr.sub<-subset(antibiotics.lr,Experiment==3|Experiment==4|Experiment==5|Experiment==6)

#Mean log reductions from data 
antibiotics.lr.stat<-antibiotics.lr.sub[,c(2:4)]
#describeBy(antibiotics.lr.sub$LR,group=list(antibiotics.lr.stat$Treatment,antibiotics.lr.stat$Time))


#Fitting a linear regression to the model 
lm.lr<-lm(LR~Treatment*Time, data=antibiotics.lr.sub)

#Check model fit 

par(mfrow=c(2,2))
plot(lm.lr)

#drop1
drop1(lm.lr,test="F")   # We see that there is a significant effect of the interaction, so we need to look at the effect of time on each treatment, below

#use lsmeans to see where the effect is


emm_table<-emmeans(lm.lr,~Treatment|Time)

#test <- pairs(contrast(emm_table))

#cld(test)

#pwpp? plot 

```


```{r, echo=TRUE}


####### LSmeans plot ##############


analysis.data<-cld(lsmeans(lm.lr,~Treatment|Time))

plot.data<-data.frame(analysis.data)
plot.data$.group[c(8,11)]<-"1,2"

#Change group numbers to letters#
plot.data$.group[plot.data$.group==" 1"]<-"a"
plot.data$.group[plot.data$.group==" 1 "]<-"a"
plot.data$.group[plot.data$.group=="  2"]<-"b"
plot.data$.group[plot.data$.group=="1,2"]<-"a,b"


error.bars<-data.frame(plot.data$upper.CL,plot.data$lower.CL)

colors2 <- c("T1" = "#56b4e9", "T2" = "#f0e442", "T3" = "#cc79a7")
facet_labels <- as_labeller(c("2" = "Time: 2 hr", "4" = "Time: 4 hr", 
                              "8" = "Time: 8 hr", "24" = "Time: 24 hr",
                              "48" = "Time: 48 hr"))
  
  
plt_lsmeans <- 
ggplot(plot.data, aes(x = Treatment, y = lsmean, color=Treatment)) +
  scale_color_manual(name="Treatment",
                       breaks=c("T1", "T2", "T3"),
                       labels=c("Tobramycin \n(100 μg/mL)", "Colistin \n(25 μg/mL)", "Tobramycin \n(100 μg/mL)\nColistin \n(25 μg/mL)"),
                       values = colors2)+
  geom_point(stat='identity',size=4) +
  scale_y_continuous(limits=c(0,10),breaks=c(0,2,4,6,8,10))+
  scale_x_discrete(breaks=c("T1","T2","T3"), labels=c("Tob","Col","Combo"))+
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.5)+
  geom_text(data=plot.data, aes(x=Treatment,y=upper.CL+0.5,label=plot.data$.group),color="black")+
  facet_grid(~Time,scales='free', labeller = facet_labels)+
  ylab("Mean Log Reduction")+
  xlab("Treatment")+
  ggtitle("Pairwise comparisons of Treatment within Time")+
  theme(axis.text.x = element_text(angle = 45,hjust=1),
        legend.direction = "horizontal",
        legend.position = c(0.28, 0.15),
        legend.box.background = element_rect(colour = "black"),
        plot.margin = margin(l=40, r = 8, t = 15),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
        text = element_text(size = 10),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.background = element_rect(fill = "#f8f8f8"),
        legend.text=element_text(size=6))

plt_lsmeans


ggsave("./figures/plt_lmeans.png", plt_lsmeans, width = 8, height = 5)
```


```{r, echo=TRUE}

### P-value plot

plt_pwpp <- 
  pwpp(emm_table)+
  scale_color_manual(name="Treatment",
                     breaks=c("T1", "T2", "T3"),
                     labels=c("Tobramycin", "Colistin", "Combination"),
                     values = colors2)+
  scale_y_discrete(labels = c("Colistin \n(25 μg/mL)",
                              "Tobramycin \n(100 μg/mL)",
                              "Tobramycin \n(100 μg/mL)\nColistin \n(25 μg/mL)"
                              ))+
  geom_vline(xintercept = 0.05, linetype="dashed", alpha = 0.5)+
  ggtitle("Pairwise P-value Plot")+
  theme(axis.text.x = element_text(angle = 60,hjust=1),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
        text = element_text(size = 10),
        plot.margin = margin(t = 15, r = 8),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.background = element_rect(fill = "#f8f8f8"),
       ) +
  facet_wrap(~Time, labeller = facet_labels, ncol = 5)
  
        
plt_pwpp 
  

ggsave("./figures/plt_pwpp.png", plt_pwpp, width = 8, height = 5)

#Combined plot 

plt_lsmeans_pwpp <- plot_grid(plt_lsmeans, plt_pwpp, nrow = 2, labels = c("A", "B"))

save_plot("./figures/plt_lsmeans_pwpp.png", plt_lsmeans_pwpp, base_height = 7, base_width = 9)

##plot showing all the points - IN PROGRESS##########


# ggplot(antibiotics.lr.sub, aes(x = Treatment, y = LR, color=Treatment)) +
#   scale_color_discrete(name="Treatment",
#                        breaks=c("T1", "T2", "T3"),
#                        labels=c("Tobramycin", "Colistin", "Combination"))+
#   theme(axis.text.x = element_text(angle = 45,hjust=1))+
#   geom_bar(stat='identity') +
#   scale_y_continuous(limits=c(0,10),breaks=c(0,2,4,6,8,10))+
#   scale_x_discrete(breaks=c("T1","T2","T3"), labels=c("Tob","Col","Combo"))+
#   geom_errorbar(antibiotics.lr.sub, width=0.2)+                  ####error bars:http://www.ats.ucla.edu/stat/r/pages/ggplot2_errorbar.htm
#   geom_text(data=plot.data, aes(x=Treatment,y=upper.CL+0.5,label=plot.data$.group),color="black")+
#   facet_grid(~Time,scales='free')+
#   ylab("Mean Log Reduction")+
#   xlab("Treatment")+
#   ggtitle("Treatment vs Time - Pairwise comparison")

```



# Antibiotic Tolerance of 24hr in New and Old Media

```{r, echo=T}

#############Statistical Analysis New vs. Old media########################## 
###############LOG REDUCTION#################################################

#Import the data 

media.lr <- read.delim("./Data/NewOldMediaAntibiotics/Log Reduction.txt")

#Change integers to factors

media.lr$Experiment<-factor(media.lr$Experiment)
media.lr$Media<-factor(media.lr$Media)
media.lr$Treatment<-factor(media.lr$Treatment)

library(plyr)
media.lr$Treatment<-revalue(media.lr$Treatment,c("1"="Tobramycin","2"="Colistin","3"="Tob/Col Combo"))


#For Experiment 1, switch Col fr and Col US values around because I must have mislabeled the flasks/plateas

media.lr[3,3]<-"Used"
media.lr[4,3]<-"Fresh"

#For experiment 3, remove Tob/col used because tech reps dont match

media.lr<-media.lr[-c(17,18),]



#Fit the model
m1<-lm(LR~Treatment*Media, data=media.lr)

#Model validation

par(mfrow=c(2,2))
plot(m1)


#Drop1 test for interaction terms

drop1(m1,test="F" )   # Shows significant interaction

#Pairwise analysis of lsmeans 
library(lsmeans)

emm_table_fresh_used<-emmeans(m1,~Media|Treatment)

pairs(contrast(emm_table_fresh_used))

cld(lsmeans(m1,~Media|Treatment))

plot(lsmeans(m1,~Media|Treatment))

#Some descriptive statistics 
library(psych)
desc.stats<-describeBy(media.lr$LR,group=list(media.lr$Treatment,media.lr$Media),mat=T)


##Plot 
library(ggplot2)

analysis.dataB<-cld(lsmeans(m1,~Media|Treatment))


plot.dataB<-data.frame(analysis.dataB)
plot.dataB$.group<-revalue(plot.dataB$.group,c(" 1 "="a","  2"="b"))


error.bars<-data.frame(plot.dataB$upper.CL,plot.dataB$lower.CL)

labels<-c("1"="Tobramycin","2"="Colistin","3"="Combination")

ggplot(plot.dataB, aes(x = Media, y = lsmean, color=Treatment)) +
  geom_point(size = 4) +
  scale_color_manual(name="Treatment",
                     breaks=c("1", "2", "3"),
                     labels=c("Tobramycin", "Colistin", "Combination"),
                     values = colors2)+
  scale_y_continuous(limits=c(-2,10),breaks=c(-2,0,2,4,6,8,10))+
  scale_x_discrete(breaks=c("Fresh","Used"), labels=c("Fresh","Used"))+
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=.1)+
  geom_text(data=plot.dataB, aes(x=Media,y=upper.CL+0.5,label=plot.dataB$.group),color="black")+
  facet_grid(~Treatment,scales='free')+
  ylab("Least-Squares Mean: Log Reduction")+
  xlab("Media")+
  ggtitle("Pairwise comparisons of Media within Treatment")+
  theme(axis.title=element_text(size=14,face="bold"),axis.text=element_text(size=14,face="bold"),
        plot.title=element_text(size=16,face="bold"))




library(ggpubr)


##All LR data 

colors3 <-  c("#56b4e9", "#f0e442", "#cc79a7")

plt_media_LR<-
ggplot(media.lr, aes(x = Media, y = LR, fill=Treatment)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(color="black", width = 0.1, height = 0, alpha = 0.5)+
  facet_grid(~Treatment)+
    scale_fill_manual(name="Treatment",
                   breaks=c("Tobramycin", "Colistin", "Tob/Col Combo"),
                   labels=c("Tobramaycin (10 μg/mL) ","Colistin (2.5 μg/mL)","Tobramycin (10 μg/mL)\nColistin (2.5 μg/mL)"),
                   values = colors3)+
  labs(y = "Log Reduction", title = "Log Reduction with Used or Fresh Media")+
    theme(axis.text.x = element_text(angle = 60,hjust=1),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
        text = element_text(size = 10),
        plot.margin = margin(t = 15, r = 8),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.background = element_rect(fill = "#f8f8f8"),
       )



plt_media_LR

ggsave("./figures/plt_media_LR.png", plt_media_LR, width = 7, height = 5)


######################################################## Log Density ###############################################################
####################################################################################################################################

#import data
media.ld <- read.delim("./Data/NewOldMediaAntibiotics/Log Density.txt")

#change integers to factors
media.ld$Experiment<-factor(media.ld$Experiment)
media.ld$Media<-factor(media.ld$Media)
media.ld$Treatment<-factor(media.ld$Treatment)

library(plyr)
media.ld$Treatment<-revalue(media.ld$Treatment,c("0"="Control","1"="Tobramycin","2"="Colistin","3"="Tob/Col Combo"))
media.ld$Media<-revalue(media.ld$Media,c("0"="Fresh","1"="Used"))

#For Experiment 1, switch Col fr and Col US values around because I must have mislabeled the flasks/plates

media.ld[5,3]<-"Used"
media.ld[6,3]<-"Fresh"


#media.lr<-media.lr[-c(17,18),] 


library(psych)
desc.stats.ld<-describeBy(media.ld$LD,group=list(media.ld$Treatment,media.ld$Media),mat=T)
desc.stats.ld

colors4 <-  c("#d55e00", "#56b4e9", "#f0e442", "#cc79a7")

###PLot
plt_media_LD<-
ggplot(media.ld, aes(x = Media, y = LD, fill=Treatment))+
  geom_boxplot(outlier.alpha = 0)+
  geom_jitter(color="black", width = 0.1, height = 0, alpha = 0.5)+
  scale_y_continuous(limits=c(0,10),breaks=c(0,2,4,6,8,10))+
  scale_fill_manual(name="Treatment",
                   breaks=c("Control","Tobramycin", "Colistin", "Tob/Col Combo"),
                   labels=c("Control","Tobramaycin (100 μg/mL) ","Colistin (25 μg/mL)","Tobramycin (100 μg/mL)\nColistin (25 μg/mL)"),
                   values = colors4)+
  ylab("Log Density (log[CFU/mL])")+
  xlab("Media")+
  ggtitle("Log Density: Media vs Treatment")+
  theme(axis.text.x = element_text(angle = 60,hjust=1),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
        text = element_text(size = 10),
        plot.margin = margin(t = 15, r = 8),
        panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
        panel.background = element_rect(fill = "#f8f8f8"),
       ) +
  facet_grid(~Treatment)

plt_media_LD


ggsave("./figures/plt_media_LD.png", plt_media_LD, height = 5, width = 7)



#Combined plot 

plt_media_combined <- 
plot_grid(plt_media_LD, 
          plt_media_LR, 
          nrow = 2, labels = c("A", "B"))

plt_media_combined

save_plot("./figures/plt_media_combined.png", plt_media_combined, base_height = 7, base_width = 9)


```

# Difference in Fluorescence Intensity (Plank-Agg)

```{r}

#import the difference data

araGFP.diff <- 
  read.csv("./Data/Plank_Agg_AraGFP/diffs.csv", header = T) 

araGFP.stats <- 
    araGFP.diff %>% 
      pivot_longer(cols = 3:8,
                   names_to = c("type", "stat"),
                   names_pattern = "(.*)_(.*)") %>%
      dplyr::select(sample, type, stat, value)

plt_araGFP_mean<-
ggplot(subset(araGFP.stats, stat == "mean"), aes(x=type, y = value))+
  geom_jitter(width=0.1, height=0)+
  geom_boxplot(width = 0.3, alpha = 0.5, outlier.alpha = 0)+
  ylim(0,8500)+
  theme(axis.text.x = element_text(angle = 60,hjust=1),
       panel.grid.minor = element_blank(),
       panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
       text = element_text(size = 10),
       plot.margin = margin(t = 15, r = 24, l = 24),
       panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
       panel.background = element_rect(fill = "#f8f8f8"),
       plot.title = element_text(size=10)
      ) +
  labs(y = "Mean Fluorescence Intensity", x = "Subpopulation", 
       title = "Mean Fluorescence Intensity \nof Single and Aggregated cells")
    
plt_araGFP_diff<-
ggplot(araGFP.diff, aes(x="", y = difference))+
  geom_jitter(width=0.05, height=0)+
  geom_boxplot(width=0.1, alpha = 0.5, outlier.alpha = 0)+
  ylim(-1000,4000)+
  theme(axis.text.x = element_text(angle = 60,hjust=1),
       panel.grid.minor = element_blank(),
       panel.grid.major = element_line(color =  rgb(235, 235, 235, 250, maxColorValue = 255)),
       text = element_text(size = 10),
       plot.margin = margin(t = 15, r = 24, b = 13, l = 24),
       panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
       panel.background = element_rect(fill = "#f8f8f8"),
       plot.title = element_text(size=10)
      ) +
  labs(y = "Difference in mean Fluorescence Intensity \n(Single Cells - Aggregates)", x = "Difference", 
     title = "Difference in Mean Fluorescence Intensity \nbetween Single and Aggregated Cells")


plt_mask_example <- 
ggdraw() + 
  draw_image("./figures/plank_agg_example.png")

bottom_row <- plot_grid(plt_araGFP_mean, plt_araGFP_diff, labels = c("B", "C"))

plt_gfp_combined <- plot_grid(plt_mask_example, bottom_row, nrow = 2, 
                              labels = c("A", NULL), rel_heights = c(0.5, 0.5), 
                              align = "hv")

save_plot("./figures/plt_gfp_combined.png", plt_gfp_combined, base_height = 7, base_width = 9)


```






